---
title: "Framingham City Data (Unofficial)"
author: "Sharon Machlis"
date: "November 9, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I will be adding information about all the files to an index.xlsx spreadsheet. index.Rmd will create an index.html file.

In general:

The data directory includes data in usable format for analysis.

The data gis subdirectory includes GIS files and maps, such as precinct/district shapefiles. 

The data elections subdirectory includes, not surprisingly, information about election results but also precinct polling places.

When initial data was not in a usable format for analysis, the original data file is in the data-raw directory. R files used to reshape the data are in the R directory. (This will not include initial conversion from PDF to Excel, which is done using the [CometDocs cloud service](https://www.cometdocs.com/).)


## Data files

*data/district_precinct_info.csv* -- which Framingham precincts make up each district.

### elections


## Raw data files

*data-raw/elections/election_general_2017_framingham.pdf* -- PDF of official Framingham November 2017 election results. [File posted on the Framingham website](http://www.framinghamma.gov/DocumentCenter/View/28924). 

*data-raw/elections/election_general_2017_framingham.xlsx* -- Official Framingham November 2017 election results converted from PDF to Excel by [CometDocs](https://www.cometdocs.com/).


## GIS files

*data/gis/FramPrecincts/* - shapefiles of Framingham precincts. Includes District info. Read in with

```{r eval=FALSE}
library(sf)
framgeo <- tmaptools::read_shape("data/gis/FramPrecincts/FramPrecincts.shp", as.sf = TRUE)

```

*data/gis/framingham_basemap.Rda* - A leaflet HTML widget of Framingham districts and precincts.

## R files

*R/config.R* -- Setup needed to run these scripts, including installing packages not on your system. **Run config.R before trying to run any other R scripts in this project.**

*R/election_general_2017_framingham.R* -- R file to convert November 2017 official results spreadsheet to an analysis-friendly format.



