---
title: "Framingham City Data (Unofficial)"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This project aims to collect important data about Framingham, Massachusetts and wrangle it so it is in a form useful for analysis. for now, much of it is election data.

Election data may be in one of two formats -- "tidy", sometimes called "long", with one observation in each row; or more human-readable format, with multiple data points per row. 

If you are doing analysis in R and other platforms (including as a basis for Excel pivot tables), tidy format is more useful. If you want to combine data sets, tidy is definitely more useful since most tidy data sets have the same column names: **Candidate, District or Precinct, Votes, Race, Year, Election**.

_Tidy election data files are in the data/elections/tidy subdirectory._

If you are just looking to see results, the non-tidy format will be more useful. Those files are in the data/elections directory.

The data generally start in 2017 with Framingham's transition from town to city form of government, although I have a few raw data files from earlier elections. I usually do not include preliminary elections unless they are city-wide. Tidy data for preliminary and general elections usually only include mayoral and City Council races. 

Source: Election data generally come from PDFs of results provided by the Framingham City Clerk's office. Note that I have then pulled that data from PDF into Excel and done a fair amount of data wrangling. 

Data added so far: 

* Some election results from the 2019 general election

* GIS shapefiles for Framigham precincts

* Results of Framingham's first-ever city general election in November 2017

* Results of Framingham's first-ever city preliminary election in September 2017


A few notes on structure and format:

* The data/gis subdirectory includes GIS files and maps, such as precinct and district shapefiles. 

* The data/elections subdirectory includes information about precinct polling places.

* When initial data was not in a usable format for analysis, such as PDFs from the City Clerk's office, the original data file is in the data-raw directory. _If you are looking for results from races not in my election CSV files, you can try looking at the raw data in the data-raw directory._

* PDFs were converted to Excel using the [CometDocs cloud service](https://www.cometdocs.com/).

* Some R files used to reshape the data have been included in the R directory. 

## Other available files

*data/district_precinct_info.csv* -- which Framingham precincts make up each district.

*data/gis/FramPrecincts/* - shapefile of Framingham precincts. Includes District info. Shapefile wsd received from Geoffrey W. Kovar, Framingham GIS Manager, via email on July 14, 2017.

Read shapefile into R with the tmaptools package:

```{r eval=FALSE}
library(sf)
framgeo <- tmaptools::read_shape("data/gis/FramPrecincts/FramPrecincts.shp", as.sf = TRUE)

```

*data/gis/framingham_basemap.Rda* - A leaflet HTML widget of Framingham districts and precincts that I created from Framingham precinct shapefile.

### R files

*R/config.R* -- Setup needed to run these scripts, including installing packages not on your system. Running config.R will ensure you have all necessary packages.

### For analysis

*R/get_turnouts.R* -- get a tidy data frame of all available election turnout data by precinct, stored in an R object called `all_turnout_data`.



## Info on reproducibility

These are the scripts I used to turn raw data into analysis-friendly data, including:

*R/election_general_2017_framingham.R* -- R file to convert November 2017 official results spreadsheet to an analysis-friendly format.

*R/election_preliminary_2017_framingham.R* -- R file to convert September 2017 official results spreadsheet to an analysis-friendly format.

*R/election_charter_2017.R* -- R file to convert April 2017 city charter results to analysis-friendly format.

*R/election_helpers.R* -- file with some useful functions for working with Framingham election data.




